---
title: "Mars_prepare"
output: html_document
---
#Set-up
```{r setup, include=FALSE}
setwd("~/DSTI/GE_project/mars-express-power-3years/mars-express-power-3years/train_set")
rm(list=ls(all=TRUE))
library(data.table)
library(zoo)
library(padr)
library(dplyr)
library(xts)
library(tabplot)
library(ggplot2)
origin="1970-01-01"
memory.limit(size=12000)
library(tidyr)
library(Matrix)
library(xgboost)
library(caTools)
library(readr)
library(stringr)
library(caret)
library(car)
```

#Prepare power
##Generate power
```{r}
power1<-read.csv("power--2008-08-22_2010-07-10.csv",sep=',',header=TRUE)
power1$ut_ms<-power1$ut_ms/1000
power1$ut_ms<-as.POSIXct(power1$ut_ms,origin=origin,tz='UTC')
power2<-read.csv("power--2010-07-10_2012-05-27.csv",sep=',',header=TRUE)
power2$ut_ms<-power2$ut_ms/1000
power2$ut_ms<-as.POSIXct(power2$ut_ms,origin=origin,tz='UTC')
power<-rbind(power1,power2)
remove(power1)
remove(power2)
power$ut_ms<-round(power$ut_ms,"secs")
#check intervals
#for(i in 1:10){
  #print(power$ut_ms[i+1]-power$ut_ms[i])}
#interval min 32s
```

###Explore power (to see if we can drop some columns)
```{r}
#check correlation
#library(corrplot)
#corrplot(cor(power[2:length(power)]))


#check columns
t<-lapply(power,mean)
z<-Reduce("+",t[2:length(t)])
t2<-t[2:length(t)]
t2<-as.data.frame(t2)
p<-prop.table(t2)
p2<-Reduce("+",p)#ok sum=1
p3<-as.data.frame(t(p))
p3$number<-colnames(p)
p4<-p3[order(-p3$V1),]
p4$cumul<-cumsum(p4$V1)#9 columns 80% power, 12 columns 90% columns
power80<-p4$number[which(p4$cumul<0.8)]
#power90<-p4$number[which(p4$cumul<0.9)]
```

###transform power(keep only variables doing 80% of the power)
```{r}
power80df<-subset(power,select=c("ut_ms",power80))
#power80df$lead<-lead(power80df$ut_ms)
#power80df$diff<-power80df$ut_ms-power80df$lead
```

###Transform in zoo.object
```{r}
power80.zoo<-read.zoo(power80df,drop="FALSE")
#remove(power)
#remove(power80df)
```

###Align power.zoo interval 60s
```{r}
xpower80<-xts(power80.zoo)
xpower80<-align.time(xpower80, n=60)
#remove(power80.zoo)
```

###Tranform in dataframe
```{r}
xpower80df<-data.frame(date = index(xpower80),xpower80,row.names = NULL)
xpower80df$NPWD2532<-as.numeric(xpower80df$NPWD2532)
xpower80df$NPWD2451<-as.numeric(xpower80df$NPWD2451)
xpower80df$NPWD2562<-as.numeric(xpower80df$NPWD2562)
xpower80df$NPWD2561<-as.numeric(xpower80df$NPWD2561)
xpower80df$NPWD2551<-as.numeric(xpower80df$NPWD2551)
xpower80df$NPWD2851<-as.numeric(xpower80df$NPWD2851)
xpower80df$NPWD2491<-as.numeric(xpower80df$NPWD2491)
xpower80df$NPWD2771<-as.numeric(xpower80df$NPWD2771)
```

###Aggregate by interval 30s
```{r}
gpower80<-xpower80df %>%
  group_by(date) %>%
  summarise_each(funs(sum(., na.rm=TRUE)))

```

```{r}
remove(p)
remove(p2)
remove(p3)
remove(p4)
remove(power)
remove(power80df)
remove(z)
remove(t)
remove(t2)
remove(xpower80)
remove(xpower80df)
remove(power80)
remove(power90)
remove(power80.zoo)
```


#Prepare saaf
###Generate saaf
```{r}
saaf1<-read.csv("context--2008-08-22_2010-07-10--saaf.csv",sep=',',header=TRUE)
saaf1$ut_ms<-saaf1$ut_ms/1000
saaf1$ut_ms<-as.POSIXct(saaf1$ut_ms,origin=origin,tz='UTC')
saaf2<-read.csv("context--2010-07-10_2012-05-27--saaf.csv",sep=',',header=TRUE)
saaf2$ut_ms<-saaf2$ut_ms/1000
saaf2$ut_ms<-as.POSIXct(saaf2$ut_ms,origin=origin,tz='UTC')
saaf<-rbind(saaf1,saaf2)
saaf$ut_ms<-round(saaf$ut_ms,"secs")
remove(saaf1)
remove(saaf2)
#check intervals
#for(i in 1:10){
  #print(saaf$ut_ms[i+1]-saaf$ut_ms[i])}
#saafcomp<-pad(saaf,start_val=minp,end_val=maxp)
#interval min 22s, around 1 min generally
```

###Transform in zoo.object
```{r}
saaf.zoo<-read.zoo(saaf,drop="FALSE")
#remove(power)
#remove(power80df)
```
###Align power.zoo interval 60s
```{r}
xsaaf<-xts(saaf.zoo)
xsaaf<-align.time(xsaaf, n=60)
#remove(power80.zoo)
```

###Tranform in dataframe
```{r}
xsaafdf<-data.frame(date = index(xsaaf),xsaaf,row.names = NULL)
```

#Aggregate by interval 60s
```{r}
gsaaf<-xsaafdf %>%
  group_by(date) %>%
  summarise_each(funs(mean(., na.rm=TRUE)))
```


```{r}
remove(saaf)
remove(xsaaf)
remove(xsaafdf)
remove(saaf.zoo)
```

#Generate FTL

```{r}
ftl1<-read.csv("context--2008-08-22_2010-07-10--ftl.csv",sep=',',header=TRUE)
ftl1$ut_ms<-ftl1$utb_ms/1000
ftl1$utb_ms<-NULL
ftl1$ute_ms<-ftl1$ute_ms/1000
ftl1$ut_ms<-as.POSIXct(ftl1$ut_ms,origin=origin,tz='UTC')
ftl1$ute_ms<-as.POSIXct(ftl1$ute_ms,origin=origin,tz='UTC')
ftl2<-read.csv("context--2010-07-10_2012-05-27--ftl.csv",sep=',',header=TRUE)
ftl2$ut_ms<-ftl2$utb_ms/1000
ftl2$utb_ms<-NULL
ftl2$ute_ms<-ftl2$ute_ms/1000
ftl2$ut_ms<-as.POSIXct(ftl2$ut_ms,origin=origin,tz='UTC')
ftl2$ute_ms<-as.POSIXct(ftl2$ute_ms,origin=origin,tz='UTC')
ftl<-rbind(ftl1,ftl2)
ftl$ut_ms<-round(ftl$ut_ms,"secs")
remove(ftl1)
remove(ftl2)

#t<-sort(ftl$ut_ms)
#s<-sort(ftl$ute_ms)
#t<-t[2:length(t)]
#s<-s[1:(length(s)-1)]
#t-s#some missing
```

###Transform ftl
```{r}
ftl$ute_ms<-NULL
ftl<-ftl[,c(3,1,2)]
```
39622 obs

###Transform in zoo.object
```{r}
ftl.zoo<-read.zoo(ftl,drop="FALSE")
#remove(power)
#remove(power80df)
```
###Align power.zoo interval 60s
```{r}
xftl<-xts(ftl.zoo)
xftl<-align.time(xftl, n=60)
#remove(power80.zoo)
```

###Tranform in dataframe
```{r}
xftldf<-data.frame(date = index(xftl),xftl,row.names = NULL)
gftldf<-xftldf
length(unique(xftldf$date))#ok no need to aggregate
gftldf<-xftldf
```

```{r}
remove(ftl)
remove(xftl)
remove(xftldf)
remove(ftl.zoo)
```

#Generate dmop
```{r}
dmop1<-read.csv("context--2008-08-22_2010-07-10--dmop.csv",sep=',',header=TRUE)
dmop1$ut_ms<-dmop1$ut_ms/1000
dmop1$ut_ms<-as.POSIXct(dmop1$ut_ms,origin=origin,tz='UTC')
dmop2<-read.csv("context--2010-07-10_2012-05-27--dmop.csv",sep=',',header=TRUE)
dmop2$ut_ms<-dmop2$ut_ms/1000
dmop2$ut_ms<-as.POSIXct(dmop2$ut_ms,origin=origin,tz='UTC')
dmop<-rbind(dmop1,dmop2)
dmop$ut_ms<-round(dmop$ut_ms,"secs")
remove(dmop1)
remove(dmop2)
t<-as.character(dmop$subsystem)
dmop$smallsub<-as.factor(substr(t,1,5))
remove(t)
```

###Transform in zoo.object
```{r}
dmop.zoo<-read.zoo(dmop,drop="FALSE")
#remove(power)
#remove(power80df)
```
###Align power.zoo interval 30s
```{r}
xdmop<-xts(dmop.zoo)
xdmop<-align.time(xdmop, n=60)
#remove(power80.zoo)
```

###Tranform in dataframe
```{r}
xdmopdf<-data.frame(date = index(xdmop),xdmop,row.names = NULL)
```

#Aggregate by interval 60s
```{r}
gdmop<-xdmopdf %>%
  group_by(date) %>%
  summarise_each(funs(paste(., collapse = ",")))
```

```{r}
remove(dmop)
remove(xdmop)
remove(xdmopdf)
remove(dmop.zoo)
```

#Generate evtf
```{r}
evtf1<-read.csv("context--2008-08-22_2010-07-10--evtf.csv",sep=',',header=TRUE)
evtf1$ut_ms<-evtf1$ut_ms/1000
evtf1$ut_ms<-as.POSIXct(evtf1$ut_ms,origin=origin,tz='UTC')
evtf2<-read.csv("context--2010-07-10_2012-05-27--evtf.csv",sep=',',header=TRUE)
evtf2$ut_ms<-evtf2$ut_ms/1000
evtf2$ut_ms<-as.POSIXct(evtf2$ut_ms,origin=origin,tz='UTC')
evtf<-rbind(evtf1,evtf2)
evtf$ut_ms<-round(evtf$ut_ms,"secs")
remove(evtf1)
remove(evtf2)
#check intervals
#for(i in 1:10){
  #print(evtf$ut_ms[i+1]-evtf$ut_ms[i])}
#min7s
```

###Transform evtf
```{r}
t<-gsub('[[:digit:]]+', '', evtf$description)
t<-gsub("*_/_",",",t)
evtf$desc2<-as.factor(t)
t<-strsplit(as.character(evtf$desc2),',')
t<-unlist(t,recursive=FALSE)
unique(t)

```

###Transform in zoo.object
```{r}
evtf.zoo<-read.zoo(evtf,drop="FALSE")
#remove(power)
#remove(power80df)
```
###Align power.zoo interval 30s
```{r}
xevtf<-xts(evtf.zoo)
xevtf<-align.time(xevtf, n=60)
#remove(power80.zoo)
```

###Tranform in dataframe
```{r}
xevtfdf<-data.frame(date = index(xevtf),xevtf,row.names = NULL)
```

#Aggregate by interval 60s
```{r}
gevtf<-xevtfdf %>%
  group_by(date) %>%
  summarise_each(funs(paste(., collapse = ",")))
```

```{r}
remove(evtf)
remove(xevtf)
remove(xevtfdf)
remove(evtf.zoo)
remove(t)
```

#Aggregate all
```{r}
gallna<-full_join(gpower80,gsaaf,by='date')
gallna<-full_join(gallna,gdmop,by='date')
gallna<-full_join(gallna,gevtf,by='date')
gallna<-full_join(gallna,gftldf,by='date')
gallna$subsystem<-as.factor(gallna$subsystem)
gallna$smallsub<-as.factor(gallna$smallsub)
gallna$description<-as.factor(gallna$description)
gallna$desc2<-as.factor(gallna$desc2)
gallna2<-subset(gallna,!is.na(gallna$NPWD2532))
summary(gallna2)
```


#Fill missing
```{r}
gall<-gallna
gall$type<- na.locf(gall$type)
gall$flagcomms<- na.locf(gall$flagcomms)
gall$sx=0.5*(na.locf(gall$sx,fromlast=TRUE) + na.locf(gall$sx))
gall$sa=0.5*(na.locf(gall$sa,fromlast=TRUE) + na.locf(gall$sa))
gall$sy=0.5*(na.locf(gall$sy,fromlast=TRUE) + na.locf(gall$sy))
gall$sz=0.5*(na.locf(gall$sz,fromlast=TRUE) + na.locf(gall$sz))
gall$desc2<- na.locf(gall$desc2,na.rm=FALSE)
gall$description<- na.locf(gall$description,na.rm=FALSE)
gall$subsystem<- na.locf(gall$subsystem,na.rm=FALSE)
gall$smallsub<- na.locf(gall$smallsub,na.rm=FALSE)
gall<-subset(gall,!is.na(gall$description))
gall<-subset(gall,!is.na(gall$desc2))
gall2<-subset(gall,!is.na(gall$NPWD2532))
```



###Data exploration
```{r}
qplot(NPWD2551,sa,data=gall)
library(plyr)
test<-subset(gall,is.na(gall$flagcomms)==FALSE&is.na(gall$NPWD2532)==FALSE)
test<-subset(test,select=c('flagcomms','NPWD2532'))
mm<-aggregate(NPWD2532~flagcomms,data=test,FUN='mean')
```


#Without na
```{r}
gNPWD2532<-subset(gall,select=c('date','NPWD2532','sa','sx','sy','sz','subsystem','smallsub','description','desc2','type','flagcomms'))
dNPWD2532<-as.data.frame(gNPWD2532)

split<-sample.split(gNPWD2532$NPWD2532,SplitRatio = 0.6)
train<-subset(gNPWD2532,split==TRUE)
test<-subset(gNPWD2532,split==FALSE)
smalltrain<-subset(train,select=-c(subsystem,description,desc2,smallsub,type))
smalltest<-subset(test,select=-c(subsystem,description,desc2,smallsub,type))
smalltrain$flagcomms<-as.numeric(smalltrain$flagcomms)
smalltest$flagcomms<-as.numeric(smalltest$flagcomms)
```

#With na
```{r}
gNPWD2532na<-subset(gallna2,select=c('date','NPWD2532','sa','sx','sy','sz','subsystem','smallsub','description','desc2','type','flagcomms'))
dNPWD2532na<-as.data.frame(gNPWD2532)

split<-sample.split(gNPWD2532na$NPWD2532,SplitRatio = 0.6)
trainna<-subset(gNPWD2532na,split==TRUE)
testna<-subset(gNPWD2532na,split==FALSE)
smalltrainna<-subset(trainna,select=-c(subsystem,description,desc2,smallsub,type))
smalltestna<-subset(testna,select=-c(subsystem,description,desc2,smallsub,type))
smalltrainna$flagcomms<-as.numeric(smalltrainna$flagcomms)
smalltestna$flagcomms<-as.numeric(smalltestna$flagcomms)
```


###Model 1 wo na
```{r}
gNPWD2532<-subset(gall,select=c('date','NPWD2532','sa','sx','sy','sz','subsystem','smallsub','description','desc2','type','flagcomms'))
dNPWD2532<-as.data.frame(gNPWD2532)

split<-sample.split(gNPWD2532$NPWD2532,SplitRatio = 0.6)
train<-subset(gNPWD2532,split==TRUE)
test<-subset(gNPWD2532,split==FALSE)

library(Matrix)

previous_na_action <- options('na.action')
options(na.action='na.pass')

sparse_train<-sparse.model.matrix(NPWD2532 ~ sa+sx+sy+sz+subsystem+smallsub+description+desc2+type+flagcomms, data = train)

options(na.action=previous_na_action$na.action)

smalltrain<-subset(train,select=-c(subsystem,description,desc2,smallsub,type))
smalltest<-subset(test,select=-c(subsystem,description,desc2,smallsub,type))
smalltrain$flagcomms<-as.numeric(smalltrain$flagcomms)
smalltest$flagcomms<-as.numeric(smalltest$flagcomms)

mat_train<-as.matrix(subset(smalltrain,select=-c(NPWD2532)))
out_train<-train$NPWD2532 == "Responder"

xgb<-xgboost(data=sparse_train,label=train$NPWD2532,nrounds=1000)

```


###Model 1 
```{r}

#library(Matrix)

previous_na_action <- options('na.action')
options(na.action='na.pass')

sparse_trainna<-sparse.model.matrix(NPWD2532 ~ sa+sx+sy+sz+subsystem+smallsub+description+desc2+type+flagcomms, data = trainna)

options(na.action=previous_na_action$na.action)



#out_train<-train$NPWD2532 == "Responder"

xgbna<-xgboost(data=sparse_trainna,label=trainna$NPWD2532,nrounds=500)

```

Score test Model 1
```{r}
previous_na_action <- options('na.action')
options(na.action='na.pass')

sparse_test<-sparse.model.matrix(NPWD2532 ~ date+sa+sx+sy+sz+subsystem+smallsub+description+desc2+type+flagcomms, data = test)

options(na.action=previous_na_action$na.action)

smalltrain<-subset(train,select=-c(subsystem,description,desc2,smallsub,type))
smalltest<-subset(test,select=-c(subsystem,description,desc2,smallsub,type))
smalltrain$flagcomms<-as.numeric(smalltrain$flagcomms)
smalltest$flagcomms<-as.numeric(smalltest$flagcomms)

y_pred <- predict(xgb, sparse_test)
length(y_pred)
length(test$NPWD2532)
RMSExgbModel1wona<-sqrt(sum(mean((y_pred-test$NPWD2532)^2)))
```

Score test Model 1 wo na
```{r}
previous_na_action <- options('na.action')
options(na.action='na.pass')

sparse_testna<-sparse.model.matrix(NPWD2532 ~ date+sa+sx+sy+sz+subsystem+smallsub+description+desc2+type+flagcomms, data = testna)

options(na.action=previous_na_action$na.action)

y_pred <- predict(xgbna, sparse_testna)
length(y_pred)
length(testna$NPWD2532)
RMSExgbModel1<-sqrt(sum(mean((y_pred-testna$NPWD2532)^2)))
```



```{r}
library(randomForest)
rf<-lm(NPWD2532 ~ .,data=smalltrain)
predrf<-predict(rf,subset(smalltest,select=-c(NPWD2532)))
error<-predrf-test$NPWD2532
RMSE<-sqrt(sum(mean(error^2)))
```

```{r}
gallna3<-gall
gallna3$description<-as.character(gallna3$description)
gallnabig<-gallna3[which(gallna3$NPWD2532>5),]

```

